run_analysis.R
========================================================

## Preliminaries
### Load relevant data to be included in dataset.  ("Inertial signal" data ignored since the raw, unprocessed data was not needed for the assignment.)

### No libraries used outside of base package.
```{r}
require(markdown)
require(knitr)

### Load data and get date/time of download.
```{r}
setwd("C:/Users/Mike/Dropbox/Coursera/CleaningData/Peer_Assignment/UCIHARDataset")
dateDownloaded <- date()   # "Sun Apr 20 19:31:20 2014"
```
## Exploratory analysis

### Read the data into R


```{r loadData}
subject_train <- read.table("./train/subject_train.txt",stringsAsFactors=FALSE) # 7352 X 1
X_train <- read.table("./train/X_train.txt",stringsAsFactors=FALSE) # 7352 X 561
y_train <- read.table("./train/y_train.txt",stringsAsFactors=FALSE) # 7352 X 1
subject_test <- read.table("./test/subject_test.txt",stringsAsFactors=FALSE) # 2947 X 1
X_test <- read.table("./test/X_test.txt",stringsAsFactors=FALSE) # 2947 X 561
y_test <- read.table("./test/y_test.txt",stringsAsFactors=FALSE) # 2947 X 1
activity_labels <- read.table("activity_labels.txt",stringsAsFactors=FALSE) # 561 X 2
features <- read.table("features.txt",stringsAsFactors=FALSE) # 561 X 2
```
------

## Exploratory analysis

### Explore dimensions and head of the data sets, and check for missing values.
```{r}
dim(X_train)
head(X_train) 
# etc.
```

## Construct the tidy datasets

### Combine training and test sets with subject vectors
```{r}
train1 <- data.frame(cbind(X_train,subject_train[,1]))
train <- data.frame(cbind(train1,y_train))
test1 <- data.frame(cbind(X_test,subject_test[,1]))
test <- data.frame(cbind(test1,y_test))
```
### Equate the names of subject vectors in training and test sets
```{r}
names(train)[562]=names(test)[562]
```
### Combine the training and test sets, use the feature vector to name the variables, and name the terminal "activity" and "subject" variables.  Change variable names to lower case.

```{r}
data <- data.frame(rbind(train,test))
colnames(data)[1:561] <- features[,2]
colnames(data)[562:563] <- c("subject","activity")
names(data) = tolower(names(data))
```

### Descriptive activity names for the activities in the data set
```{r}
data[,563][data[,563]==1] <- "WALKING"
data[,563][data[,563]==2] <- "WALKING_UPSTAIRS"
data[,563][data[,563]==3] <- "WALKING_DOWNSTAIRS"
data[,563][data[,563]==4] <- "SITTING"
data[,563][data[,563]==5] <- "STANDING"
data[,563][data[,563]==6] <- "LAYING"
```
##  "data" is a tidy data set!

### Look at structure and check for missing values
```{r}
str(data)
which(is.na(data)) # integer(0) -> no missing values

### Create a vector corresponding to variables that refer only to mean and standard deviation measurements.

```{r}
MeanStdVec = sort(union(grep("-mean()",names(data),fixed=TRUE),grep("-std()",names(data),fixed=TRUE)))
length(MeanStdVec) # a vector of length 66
```

## Combine the "-mean()/-std()" subset of "data" with the activity and subject vectors, to form "dataSub".

```{r}
dataSub = cbind(subset(data,select=MeanStdVec),data[,562:563])
```

## Create a second, independent tidy data set with the average of each variable for each activity (rows 1-6) and for each subject (rows 7-36).
```{r}
ActMeans <- as.data.frame(aggregate(dataSub[,1:67],dataSub["activity"],mean))
SubjMeans <- as.data.frame(aggregate(dataSub[,1:67],dataSub["subject"],mean))
data2 <- as.data.frame(rbind(ActMeans[,2:67],SubjMeans[,2:67]))
dim(data2) # 36 X 66
```
## Prepend "activity" and "subject" columns
```{r}
subject <- c(rep("0",6),seq(30))
activity <- c("WALKING","WALKING_UPSTAIRS","WALKING_DOWNSTAIRS","SITTING","STANDING","LAYING",rep("",30))

data2 <- as.data.frame(cbind(activity,subject,data2))
write.table(data2, file="./tidy2.txt", sep="\t", row.names=FALSE)
```






